A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Electron Beam Dynamics

### 2018-09-29 paper/theor-experiment.tex
grammer edits to theor-exp

From Table~\ref{tbl:bench_rms_error} we deduce that we achieved satisfactory
  convergence to the sampled reference Pareto front after 1000 (plus the
  additional 100 evaluations for the initial population) function evaluations.

\subsection{AWA Photoinjector Optimization} \label{awaproblem}
Next we apply the optimization framework to a high charge beam line
 at the Argonne Wakefield Accelerator (AWA) facility. 
The goal of this optimization is to produce beams of electrons that meet 
design specifications; this includes number of particles (charge), energy, and particle distribution (characterized by beam sizes and energy spread).
As shown in Fig.~\ref{awa-linac}, the installed portion of the the 
beam line consists of an rf photocathode gun, 
two solenoids, and six linear accelerating cavities
followed by four quadrupoles and a stripline kicker. 
The charge of interest, 40 nC, is needed for Two Beam 
Acceleration (TBA) 
experiments performed at AWA \cite{gai_power_jing_2012,JING201872}, 
which motivates this work. 
Prior experimental results were limited by beam size when the beam passed through small aperture 
wakefield structures located downstream. 
In an attempt to maximize charge transmission in upcoming experiments
we optimize magnet strengths of the solenoids and quadrupoles leading into the 
TBA section of the beam line, shown in Fig.~\ref{awa-tba}.
The simulation model includes from the gun to the septum.
The optimization location is chosen as the  
entrance to the first quadrupole on the dog leg ($s_3$), see Fig.~\ref{awa-tba}.
Minimizing beam sizes here will enable capture and further focusing before space charge effects dominate the beam. 
This will also enable cleaner transport through downstream elements.
%\vspace{-1em}

@@

In addition to addressing the challenge of producing an optimized beam, 
we chose this model to demonstrate the ability of the framework to tackle large problems.
Six design variables and objectives were used, along with three constraints.
The objectives include transverse and longitudinal beam sizes, 
transverse momentum, and longitudinal energy spread. 
The design variables include the two gun solenoids and 
the first four quadrupoles settings. 
This problem encompasses high dimensionality 
and nonlinear effects such as space charge. There is no existing information
that accurately predicts optimized parameters for this beam line. 
This work is meaningful in that it will guide future operations at the AWA.

\subsubsection{Time Step Scan} \label{awa:subsection:test}
Before running a full scale optimization of the problem described in Subsection \ref{awaproblem}, 
a study on time step and number of particles in the simulation model 
was done to reduce the time of the simulation while 
maintaining the physics of interest. 
The grid size $16 \times 16 \times 32$ was chosen, 
and parallelized in the x and y directions.
After comparing several options (1,000, 10,000, 20,000, 50,000, 100,000) 
with a small time step, the number of particles was fixed at ten thousand.
Next several time steps were explored, see Table~\ref{timestep}.
The largest steps were too big to resolve the beam parameters accurately.
See low fidelity plot in Fig.~\ref{tstep} for $dT=$\num{5E-11} results.

@@

In the drifts and linac tanks, $dT=\num{1E-11}$ was sufficient, 
but not acceptable near the quadrupoles. 
For all models, the longitudinal parameters (rms$_s$ and energy) 
are calculated correctly, but we see discrepancies in the transverse 
(rms$_x$ and $\epsilon_x$) for low fidelity results. This discrepancy is 
what led to the decision to adjust the time steps w.r.t. beam line elements. 
In the linac and drift sections $dT=\num{1E-11}$ was used. 
A time step of $dT=\num{1E-12}$ was used near sensitive elements such as the quadrupoles, kicker, and septum.

>>>>

Table~\ref{tbl:bench_rms_error} shows satisfactory
  convergence to the sampled reference Pareto front after 1000 (plus the
  additional 100 evaluations for the initial population) function evaluations.

\subsection{AWA Photoinjector Optimization} \label{awaproblem}
Next the optimization framework is applied to a high charge beam line
 at the Argonne Wakefield Accelerator (AWA) facility. 
The goal of this optimization is to produce beams of electrons that meet 
design specifications; this includes number of particles (charge), energy, and particle distribution (characterized by beam sizes and energy spread).
As shown in Fig.~\ref{awa-linac}, the installed portion of the the 
beam line consists of an rf photocathode gun, 
two solenoids, and six linear accelerating cavities
followed by four quadrupoles and a stripline kicker. 
The charge of interest, 40 nC, is needed for Two Beam 
Acceleration (TBA) 
experiments performed at AWA \cite{gai_power_jing_2012,JING201872}, 
which motivates this work. 
Prior experimental results were limited by beam size when the beam passed through small aperture 
wakefield structures located downstream. 
In an attempt to maximize charge transmission in upcoming experiments,
magnet strengths of the solenoids and quadrupoles leading into the 
TBA section of the beam line were optimized, shown in Fig.~\ref{awa-tba}.
The simulation model includes from the gun to the septum.
The optimization location is chosen as the  
entrance to the first quadrupole on the dog leg ($s_3$), see Fig.~\ref{awa-tba}.
Minimizing beam sizes here will enable capture and further focusing before space charge effects dominate the beam. 
This will also enable cleaner transport through downstream elements.
%\vspace{-1em}

@@

In addition to addressing the challenge of producing an optimized beam, 
this model was chosen to demonstrate the ability of the framework to tackle large problems.
Six design variables and objectives were used, along with three constraints.
The objectives include transverse and longitudinal beam sizes, 
transverse momentum, and longitudinal energy spread. 
The design variables include the two gun solenoids and 
the first four quadrupoles strengths. 
This problem encompasses high dimensionality 
and nonlinear effects such as space charge. There is no existing information
that accurately predicts optimized parameters for this beam line. 
This work is meaningful in that it will guide future operations at the AWA.

\subsubsection{Time Step Scan} \label{awa:subsection:test}
Before running a full scale optimization of the problem described in Subsection \ref{awaproblem}, 
a study on time step and number of particles in the simulation model 
was done to reduce the time of the simulation while 
maintaining the physics of interest. 
The grid size $16 \times 16 \times 32$ was chosen, 
and parallelized in the x and y directions.
After comparing several options (1,000, 10,000, 20,000, 50,000, 100,000) 
with a small time step, the number of particles was fixed at 10,000.
Next several time steps were explored, see Table~\ref{timestep}.
The largest steps were too big to resolve the beam parameters accurately.
See low fidelity plot in Fig.~\ref{tstep} for $dT=$\num{5E-11} results.

@@

In the drifts and linac tanks, $dT=\num{1E-11}$ was sufficient. 
However, it was not acceptable near the quadrupoles. 
For all models, the longitudinal parameters (rms$_s$ and energy) 
are calculated correctly, but we see discrepancies in the transverse 
(rms$_x$ and $\epsilon_x$) for low fidelity results. This discrepancy is 
what led to the decision to adjust the time steps w.r.t. beam line elements. 
In the linac and drift sections $dT=\num{1E-11}$ was used. 
A time step of $dT=\num{1E-12}$ was used near sensitive elements such as the quadrupoles, kicker, and septum.
